---
title: "Solar Energy"
subtitle: "Data Visualization 608"
author: "Maria A Ginorio"
date: "11/12/2022"
output: 
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: true
---

## Load Packages

As usual, it is first necessary to load some packages before building the figure. Today's chart is going to use the help of ggrepel and ggtext. ggrepel will make it extremely easy to automatically position labels and ggtext is going to make it possible to use annotations that mix normal text with italic and bold text

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = T)

# Data analysis libraries
library(tidyverse)
library(magrittr) # extended pipe functionality
library(lubridate) # dealing with date objects
library(readxl) # for reading Excel files
library(skimr)
```

```{r packages, warning=FALSE, message=FALSE}

# Visualization libraries
library(ggrepel) # for neater text labels
library(gghighlight) # for easier highlighting
library(extrafont) # for... extra fonts
library(gridExtra) # for combining plots
library(geofacet) # for making facets 
library(ggridges)
library(ggtext)
library(showtext)

font_add_google("Lato")
showtext_auto()

```

## Load & Prepare Dataset

First, we'll take a look at data on solar photovoltaic energy consumed by the commercial, industrial, power and residential sectors, courtesy of the U.S. Energy Information Administration's [State Energy Data Systems (SEDS) database](https://www.eia.gov/state/seds/).

We start by utilizing the data sets obtained by SEDS (U.S Energy Information Administration) [SEDS](https://www.eia.gov/renewable/data.php). We will utilize 3 datasets related to the Renewable Energy Consumption by sector section.

-   Dataset 1: Residential

-   Dataset 3: Commercial

-   Dataset 3: Industrial

Then we will proceed to change the format of the data to a longer format using the tidyverse package:

```{r dataset , warning=FALSE, message=FALSE}
processed_renewals_energy <- read_csv("processed_renewals_energy.csv")
```

```{r}
# df1 <- Processed_Renewable_Sector %>% 
#   rename(Year = `Annual Total`) %>% 
#   mutate(across(everything(), na_if, "Not Available")) %>% 
#   mutate(across(everything(), na_if, "No Data Reported"))

# df1 <- df1 %>% 
#   pivot_longer(
#     cols = starts_with("Geothermal"),
#     names_to = "Geothermal",
#     values_to = "BTUs",
#     names_prefix = "Geothermal",
#     values_drop_na = F 
#   )

# df1 <- df1 %>% 
#   pivot_longer(
#     cols = starts_with("Wind"),
#     names_to = "Wind",
#     values_to = "BTUs_3",
#     names_prefix = "Wind",
#     values_drop_na = F 
#   )

# df1 <- df1 %>% 
#   mutate(across(c(`Hydroelectric Commercial`), as.numeric)) %>% 
#   pivot_longer(
#     cols = starts_with("Hydroelectric"),
#     names_to = "Hydroelectric",
#     values_to = "BTUs_4",
#     names_prefix = "Hydroelectric",
#     values_drop_na = F 
#   )

# df2 <- df1 %>% 
#   pivot_longer(
#     cols = c("Hydroelectric","Wind","Solar", "Geothermal" ),
#     names_to = "energy_type",
#     values_to = "source",
#     values_drop_na = F 
#   )

# df2 <- df1 %>% 
#   pivot_longer(
#     cols = c("Hydroelectric","Wind","Solar", "Geothermal" ),
#     names_to = "energy_type",
#     values_to = "source",
#     values_drop_na = F 
#   )

# df2 <- df2 %>% 
#   mutate(across(c(BTUs, BTUs_2, BTUs_3), as.numeric)) %>% 
#   pivot_longer(
#     cols = BTUs:BTUs_4,
#     names_to = "data_no",
#     values_to = "data",
#     values_drop_na = F 
#   )

# pgen<- df2 %>% 
#   select(-data_no, -`Wood Residential`) %>% 
#   mutate(data = replace_na(data, 0))

#write_csv(pgen, "processed_renewals_energy.csv")

```

### Preprocessed Data

```{r, message=FALSE, warning=FALSE}
#write_csv(gen2, "gen2.csv")
processed_data <- read_csv("gen2.csv")
head(processed_data)
```

Next is Important to determine the highlights for our variables so we can obtain the information we are looking for. Its important that our plot shows the sources that correspond to the category solar. We make the modifications using dplyr package.

```{r highlights, warning=FALSE}
highlights <- c("Solar, Residential", "Solar, Commercial", "Solar, Industrial")
```

```{r gen2, warning=FALSE, message=FALSE}

gen2 <- processed_renewals_energy %>% 
  group_by(Year, source, energy_type) %>%
  summarize(total_btu = sum(data)) %>% 
  mutate(full_source = paste0(energy_type, ", ", source)) %>%
  mutate(
    highlight_col = ifelse(full_source %in% 
                             highlights, full_source, "other"),
    highlight_col = as.factor(highlight_col)) %>% 
  mutate(
    highlight_col = fct_relevel(highlight_col, "other", after = Inf)) %>% 
  ungroup()
```

```{r namelab, message=FALSE, warning=FALSE}
gen2 <- gen2 %>% 
  group_by(Year, highlight_col) %>% 
  mutate(
    name_lab = if_else(
      highlight_col %in% highlights, highlight_col, factor(NA))
  ) %>% 
ungroup()
```

## Theme Definition

In this occasion, the theme is defined before creating the plot. the `element_markdown()` in `plot.subtitle()` above? That function comes with the `ggtext()` package and makes it possible to use markdown syntax to format the text. It's like magic!

```{r theme2}

# This theme extends the 'theme_minimal' that comes with ggplot2.
# The "Lato" font is used as the base font. This is similar
# to the original font in Cedric's work, Avenir Next Condensed.
theme_set(theme_minimal(base_family = "Lato"))


theme_update(
  # Remove title for both x and y axes
  axis.title = element_text(),
  # Axes labels are grey
  axis.text = element_text(color = "grey40"),
  # The size of the axes labels are different for x and y.
  axis.text.x = element_text(size = 20, margin = margin(t = 5)),
  axis.text.y = element_text(size = 17, margin = margin(r = 5)),
  # Also, the ticks have a very light grey color
  axis.ticks = element_line(color = "grey91", linewidth = .5),
  # The length of the axis ticks is increased.
  axis.ticks.length.x = unit(1.3, "lines"),
  axis.ticks.length.y = unit(.7, "lines"),
  # Remove the grid lines that come with ggplot2 plots by default
  panel.grid = element_blank(),
  # Customize margin values (top, right, bottom, left)
  plot.margin = margin(20, 40, 20, 40),
  # Use a light grey color for the background of both the plot and the panel
  plot.background = element_rect(fill = "grey98", color = "grey98"),
  panel.background = element_rect(fill = "grey98", color = "grey98"),
  # Customize title appearence
  plot.title = element_text(
    color = "grey10", 
    size = 20, 
    face = "bold",
    margin = margin(t = 15)
  ),
  # Customize subtitle appearence
  plot.subtitle = element_markdown(
    color = "grey30", 
    size = 16,
    lineheight = 1.35,
    margin = margin(t = 15, b = 40)
  ),
  # Title and caption are going to be aligned
  plot.title.position = "plot",
  plot.caption.position = "plot",
  plot.caption = element_text(
    color = "grey30", 
    size = 13,
    lineheight = 1.2, 
    hjust = 0,
    margin = margin(t = 40) # Large margin on the top of the caption.
  ),
  # Remove legend
  legend.position = "right"
)

```

## Basic Chart

Today's chart is a lineplot that visualizes the consumption changes (in BTU's) of a Solar Energy based on a time frame from 1960 - 2000. The full_source in the vector `highlights` have different colors to stand out from the rest and make it easier to track their evolution along time.

```{r plt1}
plt <- ggplot(
  gen2 %>% filter(highlight_col != "other"),
  aes(Year, total_btu, 
             group = full_source)) + 
  
  # Geometric annotations that play the role of grid lines
  geom_vline(
    xintercept = seq(1960, 2020, by = 5),
    color = "grey91", 
    linewidth = .6
  ) + 
  geom_segment(
    data = tibble(y = seq(0, 100000, by = 10), x1 = 1960, x2 = 2020),
    aes(x = x1, xend = x2, y = y, yend = y),
    inherit.aes = FALSE,
    color = "grey91",
    linewidth = .6
    #change log 10 for by 100 if it doenst work
  ) +
  geom_line(
    data = gen2 %>% filter(highlight_col == "other"),
    color = "grey75",
    linewidth = .6,
    alpha = .5,
   ) + 
  ## Lines for the highlighted countries.
  # It's important to put them after the grey lines
  # so the colored ones are on top
  geom_line(
    aes(color = highlight_col ),
    linewidth = .9,
    na.rm = F,
    orientation = NA,
  )
  
```

```{r}
plt
```

## Add Labels

Although colors are a tremendous help, they don't reveal which country the lines represent. Wouldn't it be nice to have a label on the end of each line that tells which country it represents?

It is pretty challenging to add many labels on a plot since labels tend to overlap each other, making the figure unreadable. Fortunately, the `ggrepel` package is here to help us. It provides an algorithm that will automatically place the labels for us. Let's do it!

```{r, warning=FALSE, message=FALSE}
plt <- plt + 
  geom_text_repel(
    aes(color = highlight_col, label = name_lab), data = gen2,
    family = "Lato",
    fontface = "bold",
    size = 15,
    direction = "y",
    xlim = c(2019, NA),
    hjust = 0,
    segment.size = .7,
    segment.alpha = .5,
    segment.linetype = "dotted",
    box.padding = .4,
    segment.curvature = -0.1,
    segment.ncp = 3,
    segment.angle = 20,
    show.legend = F
  ) + 
  ## coordinate system + scales
  coord_cartesian(
    clip = "off",
    ylim = c(10, 200000)
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    limits = c(1960, 2023.5), 
    breaks = seq(1960, 2020, by = 10)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    breaks = c(1000, 10000, 100000, 200000),
    labels = c("1K","10K","100K", "200K")
  )

#ggsave("solar_consumption.png")
```

```{r plt2, warning=FALSE, message=FALSE}
plt
```

## Solar Light Consumption

The use of renewable energy has been increasing over the past few years as the cost of traditional energy sources has risen and the environmental concerns about the use of fossil fuels have grown. Renewable energy is energy that comes from natural sources that are replenished on a human timescale, such as solar, wind, water, and geothermal.

There are many reasons why the use of renewable energy is growing. One reason is that renewable energy is becoming more cost-competitive with traditional energy sources. The cost of renewable energy technologies has fallen dramatically in recent years, and is expected to continue to fall. The cost of solar photovoltaic (PV) panels has fallen by around 90% since 2010, and the cost of wind turbines has fallen by around 50% since the early 2000s. As the cost of renewable energy falls, the cost of traditional energy sources is rising. This is due to the declining quality of traditional energy sources and the rising cost of environmental regulations.

Another reason for the growth of renewable energy is the increasing concern about the environmental impacts of fossil fuels. The burning of fossil fuels releases greenhouse gases, which are causing the Earth's climate to change. This is leading to more extreme weather events, such as heatwaves, droughts, and floods. The use of renewable energy can help to reduce the emissions of greenhouse gases and slow the rate of climate change.

The growth of renewable energy is also being driven by government policies. Many governments are setting targets for the use of renewable energy, and providing financial incentives, such as subsidies and tax breaks, to encourage the use of renewable energy. The use of renewable energy is expected to continue to grow in the future. This is due to the falling cost of renewable energy, the increasing concerns about the environment, and the support of government policies.

```{r plt3}
plt <- plt + 
  labs(
    title = "What has been the evolution\nof Solar Energy Consumption since 1960",
    subtitle = "Solar power is more affordable,accessible, and prevalent in the United States than ever before",
    caption = "Visualization by Maria A Ginorio \n • Data source: SEDS (U.S. Energy Information Administration) ")
```

```{r, message=FALSE, warning=FALSE}
plt
```
